Measuring the Influence of Twitter Bots during the 2018 US Midterm Election

### 2019-09-15 paper.tex
Latest Submission

\begin{abstract}
	This study uses 6.48M tweets and 1.6M profile data collected from Twitter using 109 keywords during the 2018 US Midterm Elections to
    determine the influence bots had. Data was collected between October 31, 2018 (1 AM UTC) and November 7, 2018 (12 AM UTC). 
    Machine Learning techniques were used on this data to separate bots and humans. Afterward, the analysis found that 1\% of the
    top bots influenced the creation of 11\% of the tweets. Additionally, for URLs which occurred
    more than ten times in our dataset, the total influence percentage of bots was double their population percentage. Here, influence is a number given to each tweet which measures 
    "the expected number of time the tweet is retweeted – direct retweets or descendants in a diffusion scenario – over all possible diffusion scenarios  associated." \cite{rizoiu2018debatenight} 
    By this definition, a tweet T\textsubscript{2} is said to be influenced by a tweet T\textsubscript{1} if T\textsubscript{2} is a retweet of T\textsubscript{1}. A tweet can be 
    influenced by multiple tweets. In the previous example, if T\textsubscript{3} retweets the retweet made by T\textsubscript{2}, it is influenced by both T\textsubscript{1} and T\textsubscript{2}.
    Influence percentage is calculated from the percentage of influence a user had calculated on aggregate from the complete dataset.
\end{abstract}

@@

Twitter is a social media where short messages get exchanged. Messages shared by a user are shown to
their followers and to followers of those who share the message. These messages are called Tweets. Sharing can be done by retweeting
(sharing) or replying to the tweet. When Alice replies/retweets, a tweet made by Bob, Alice and Bob's followers will see the tweet. In this
way, messages spread. Having massive followers, getting many retweets or getting retweeted by people with massive followers can make a tweet viral.\par

People use social media, at least in part, to form an opinion about lifestyle, health, politics, and purchase \cite{varol2017early}. Cost of advertising on Twitter 
can reflect the value of a Tweet. In 2018 Q4 the median Cost Per Mile (CPM) - the cost for reaching 1000 people
was - \$5.93 while the median Cost Per Click (CPC) was \$0.4. The median Click Through Rate (CTR) was 1.55\% \cite{adstage_2018_report}. 
Due to the associated cost and possible underlying problems, advertising is not always feasible. 
Various other techniques have been devised and are also combined with advertisements to reach a big audience. One such technique is the use of bots. 
Social bots are account controlled wholly or partially by a computer algorithm. 
These bots can generate content and interact with human users often imitating humans \cite{ferrara2016rise}.
 A US special counsel investigation found that Internet Research Agency (IRA), a St Petersberg based company,
created Twitter accounts of US personas and operated a botnet (a network of bots) to amplify content. 
During the US Presidential Election, they explicitly supported or opposed political candidate \cite{luceri2019red,mueller_investigation}. 
These propaganda bots were found to be high volume, multichannel, rapid, continuous, repetitive, inaccurate, and inconsistent \cite{paul2016russian}. \par

Studies have explored political polarization \cite{munger2017don,rizoiu2018debatenight,gruzd2014investigating,bovet2019influence} , origin of bots 
\cite{zannettou2019let,zannettou2019characterizing} and the spread of fake news \cite{vosoughi2018spread,shao2018spread,lazer2018science,bovet2019influence,grinberg2019fake}
on twitter. Most studies focus on 2016 Election \cite{bovet2019influence,rizoiu2018debatenight,bessi2016social,howard2018algorithms,howard2016bots}. One studied the 2018 Midterm 
\cite{deb2019perils}. Few studies perform a large scale quantitative study of the role played by bots during a significant event. Using ML techniques to detect bots and measure their success/failure during a significant event, 
this study attempts to add to the small body of literature. \par

@@

\begin{itemize}
    \item 10.5\% of the users in the datasets were bots. They made 12.6\% of the tweets and created 22\% of the total influence.
    \item On Average, bots were twice as influential as humans
    \item Top 1\% of the bots influenced the creation of 11\% of retweets in the dataset
    \item In average a single URL was shared 27.84 times in 1.97 cascades
    \item The influence percentage of bots in URL shared was similar to their population. However, for URLs which were shared more than ten times, their influence percentage was twice their population.
    \item Bots dominate humans in the use of conservative words. Humans dominate bots in the use of slangs like yall, bc, and bunch.
    \item Tweets made by bots was retweeted by bots 2.87\% or 120k times. Tweets made by bots were retweeted 
    by humans 6.34\% or 265k times. Tweets made by humans were retweeted by bots 9.14\% or 382k times. Finally, most of the interaction was between humans and humans i.e., 81.64\% or 
    3.4M times.
    \item Most of the codes and notebooks are shared publicly in https://github.com/warproxxx/2018Midterm to help future researchers and to ensures transparency.
\end{itemize}

@@

Different types of bots exist in social media. A simple bot will post predetermined messages at predetermined intervals \cite{haustein2016tweets}. Bots are also used to increase followers
and make an account appear popular \cite{cresci2015fame}. Botnets are a network of bots working together. The word Botnet is also used to refers to a network of compromised computer. But, in this report, 
the word botnet refers to a network of social media bots working together to make an influence and should not be confused with the other definition. Many botnets use a hybrid human/automation 
approach \cite{grimme2018changing}. Botnets have been used to promote spam \cite{ferrara2018measuring} (which seem to shifted to social media due to the effectiveness of spam filter \cite{gao2010detecting,chu2012detecting,ferrara2018measuring}), 
manipulate stocks \cite{ferrara2015manipulation}, manipulate elections \cite{morstatter2018alt}, 
and numerous other purposes \cite{abokhodair2015dissecting}.
Botnets were detected as early as 2010 US Midterm Election \cite{mitter2014categorization}. Over time, their use and study have increased. 
Studies have detected and analyzed bots in the 2017 German Federal Election \cite{morstatter2018alt} ,2017 French Presidential Election \cite{ferrara2017disinformation} and during 
various US Elections \cite{mitter2014categorization,bovet2019influence,rizoiu2018debatenight,bessi2016social,howard2018algorithms,howard2016bots,deb2019perils}. In Elections, 
they have been used to support candidates \cite{luceri2019red}, attack people \cite{mueller_investigation} and spread fake news \cite{vosoughi2018spread,grinberg2019fake}. 
Political bots have been used beyond elections. \cite{stewart2018examining} found that Russian bots infiltrated both right and left-leaning communities and spread different narratives 
\cite{mueller_investigation}. \par

In Twitter, bots share content 
in multiple channel \cite{paul2016russian}. This activity is in line with literature which shows that information from multiple sources appears more trustworthy \cite{harkins1981multiple}. 
 Bots attack people. Attacking trustworthiness has been shown to diminish the credibility of the Original Poster \cite{pornpitakpan2004persuasiveness}. 
Bots are continuous, repetitive, but most of the times, they are also far from reality \cite{paul2016russian}. The high volume seems to ensure early exposure and \cite{petty1994think} shows that early
 first impression is more likely to be accepted by the brain. People have confirmation bias reinforcing their views. However, low credibility
  sources can have persuasive power even to unbiased or neutral people. When remembering, information gets disassociated from the source in a phenomenon known as sleeper effect \cite{underwood1998memory,paul2016russian}. 
As a boost to sleeper effect, people are 31\% more likely to remember what they see on twitter, compared to the normal web \cite{twitter_remember}. \par

\subsection{Bot Detection}
Botometer, the only openly accessible bot detection system used in academia \cite{davis2016botornot,yang2019arming}, has been widely used \cite{rizoiu2018debatenight,yang2019arming,shao2018spread}.
Botometer's Machine Learning algorithm provides an account-level bot classification. Botometer classifies account partially and completely automatized as bots. It has some false positives 
, and most notably, its detection of organizational accounts as bots has been highly criticized \cite{varol2017early} \cite{botometer_tweet}. However, even with its false positive, 
Botometer has no transparent competitor and 
remains the most accurate bot detection system in academia. \par

@@

\cite{rizoiu2018debatenight} found that social bots were 2.5 times more influential than humans. They introduce a scalable algorithm for estimating user influence in retweet cascades. 
For each tweet in an information cascade, it uses time of post and the number of followers to determine the probability of whether a tweet is a retweet of another. They then find influence which can be used 
to find out how many users the tweet possibly influenced to retweet. It was tested successfully in an artificial social media. Other studies have attempted to measure influence before this. 
\cite{weng2010twitterrank} used eigenvector centrality of the connection to measure influence, but it is not scalable into big cascades. There are other methods like 
\cite{rodriguez2011uncovering,cho2013latent,linderman2014discovering} but they either have scaling issue or require full diffusion graph, which Twitter does not provide.

\section{Methods}
\label{sec:method}
\subsection{Data Collection}
Twenty-nine manually identified keywords, and names of all 53 House candidates, and 27 Senate candidates were used to collect live data from Twitter. The manually defined keywords were:

@@

Twitter provides Streaming API, Search API, and a premium Firehose API to provide access to tweets. Streaming API provides limited access to live data (termed 1\% API in the early days) while Firehose API
provides complete data. However, due to the associated costs, Firehose API was not an option in this study. \cite{morstatter2013sample} compares the Streaming API with Firehose API.
They found that Streaming API was nearly as good as a random sample of Firehose API when the dataset was large enough. Although the 1\% API was not as good as a 1\% random sample from the Firehose API in 
all of their tests, it estimated the top hashtags correctly when the data was large enough. It recommended the creation of specific parameters and using a large sample. 
Due to the possible issues in Streaming API, \cite{bessi2016social} recommends using Twitter Search API.

\par This study uses the Streaming API with many keywords aiming to collect a large dataset. While collecting data, information about the tweeter and the tweet was collected. 
Collected information about a tweet is:
\begin{enumerate}[label=\textbf{\arabic*}]
    \item \textbf{Timestamp:} UTC Timestamp in which the post was made
    \item \textbf{ID:} Post ID provided by twitter using their snowflake algorithm
    \item \textbf{Text:} Text in the tweet and the parent tweet if the tweet is a reply
    \item \textbf{User:} Username of the tweeter/retweeter/replier
    \item \textbf{Replies:} The number of replies the tweet has received. During live collection, this value is zero.
    \item \textbf{Retweets:} The number of retweets the tweet has received. During live collection, this value is zero.
    \item \textbf{Likes:} The number of likes the tweet has received. During live collection, this value is zero.
    \item \textbf{Reply To ID:} The ID of the parents tweet if this is retweet or a reply.
    \item \textbf{Response Type:} Either Tweet or Retweet or Reply
\end{enumerate}
\bigskip
Profile data were collected along with the tweets

\begin{enumerate}[label=\textbf{\arabic*}]
    \item \textbf{Username:} Username of the user who posted the tweet
    \item \textbf{Location:} Binary if geolocation is enabled
    \item \textbf{Is Verified:} A binary if the profile has been verified
    \item \textbf{Total Tweets:} Total number of tweets created by the user
    \item \textbf{Total Following:} Total accounts following
    \item \textbf{Total Followers:} Total Followers
    \item \textbf{Total Listed:} Number of times a user has been listed
    \item \textbf{Total Status:} Total Status
    \item \textbf{Total Likes:} Total likes the user has received
    \item \textbf{Has Background:}  Binary if an account has a background
    \item \textbf{Is Protected:}  Binary if an account is protected
    \item \textbf{Profile Modified:}  Binary if a profile has been modified
\end{enumerate}
\bigskip
These parameters are same as the ones used by \cite{kudugunta2018deep} and \cite{ferrara2017disinformation}. Same data was collected because the Machine Learning architecture used for bot 
detection is same. \par

These parameters were used to collect Twitter data from October 31, 2018 (1 AM UTC) to November 7, 2018 (12 AM UTC). Twitter API was used to re-scrape the details and update the number of likes, replies, and retweets of top 100 thousand tweets  
in the dataset on July 18, 2019.

\subsection{Data Processing}
First, irrelevant tweets were manually removed. VOTE was one of the keywords, and other people, most notable of which were the followers of music bands, had used it too. Keywords were created to find and remove these irrelevant data.\par

SentiStrength and VADER Sentiment analysis was used. SentiStrength \cite{thelwall2010sentiment} is used to annotate short, informal tweet like texts \cite{bessi2016social}. 
It can capture positive and negative emotions at an accuracy of 60.6\% and 72.8\%. SentiStrength gives a positive and negative score between 0 and 4. The positive sentiment is subtracted 
from the negative like in \cite{bessi2016social} to get a whole number. VADER is a rule-based sentiment classification designed for use in social media data. VADER was found to outperform 
individual human raters with an F1 accuracy of 0.96 compared to 0.84 for humans \cite{hutto2014vader}.

@@

\begin{itemize}
    \item \cite{cresci2017paradigm} provided a labeled account level bot or not data with relevant features that included the ones we collected.
    These features were used to train a Random Forest Classifier and detect bots. 
    \cite{cresci2017paradigm}'s data is old, and bots evolve daily. So further processing was required.
    \item The trained model was used in our dataset, which contains the same features, to select 40k highly probable bots.
    \item 40k highly probable bots and 32k random accounts were selected, and bot detection was done through Botometer. Botometer returns CAP score and bot score. CAP score denotes the probability of being a bot. 
    Studies have used bot score of 0.5 to determine bots. CAP score of 0.3 is equivalent to the score 0.5 used by researchers before. \cite{deb2019perils} used a threshold of 0.3 after the 
    introduction of new changes as mentioned in \cite{yang2019arming}.
    \item Grid search was performed with thresholds of 0.1, 0.2, and 0.3 for humans, 0.3, 0.5, 0.7, and 0.8 for bots and Random Forest, Neural Network and SMOTENN as possible ML algorithms.
    \item It was determined that 0.5 was the optimal bot threshold, 0.3 the optimal human threshold and SMOTENN the optimal ML algorithm. Using this threshold, we selected 513 bots whose CAP score was higher than 0.5 , 
    +27k clean accounts whose CAP score was smaller than 0.3 for training and testing our data. The remaining 44k values that lied between these were used as a broader test set.
\end{itemize}

We obtained an accuracy of 85\% with AUC of 0.85 in the training set, 87\% accuracy with 
an AUC of 0.8 in the smaller test set and accuracy of 85\% with an AUC of 0.8 in the broader test set when compared with Botometers prediction. This model detected 10.5\% of the accounts as 
bots. The accuracy was lesser than in \cite{kudugunta2018deep} likely due to the evolution of bots or because we compared with Botometer instead of manually labeling like in the study.

\subsection{Cascading and Influence Detection}
A cascade starts when a Tweet is made. Any retweet or response which involve that Tweet belongs to that cascade. If Bob retweets Alice's tweet, and Eve retweets Bob's retweet, the tweets
made by Alice, Bob, and Eve will belong in the same cascade. The Twitter API will show Eve and Bob as a direct descendant of Alice without any mention that Eve retweeted Bob's retweet. \par

\cite{du2013scalable} defines influence as the average number of users who get in contact with the content created by a user \textit{u}.
 However, the number of people reached, and the diffusion graph is not provided by Twitter. \cite{rizoiu2018debatenight} define the influence of a user over a retweet cascade as "the expected number of time the tweet is
 retweeted – direct retweets or descendants in a diffusion scenario – over all possible diffusion scenarios  associated." Influence of a user is defined as the sum of the influence of tweets
 authored by a user. As their algorithm is used, this definition is used. \cite{rizoiu2018debatenight} uses time and 
 number of followers to estimate a user influence. Using this mechanism, a high influence score can be provided to highly connected users who never start diffusions and 
 to active retweeters with little followership. \cite{rizoiu2018debatenight} tested their algorithm on an artificial social network with 1000 users and found that the influence calculated had a Spearman 
 correlation coefficient of 0.88 with the actual influence. \par

@@

 \cite{rizoiu2018debatenight} tested the validation of their algorithm when they had access to full data. We did not. So random cascades were synthesized using the following algorithm 
 to test the validity:

@@

 The influence was then converted to a user wise percentage, and correlation was calculation between the actual and predicted value. A high correlation was observed. 
 This is a duct tape algorithm and should be treated as such. It would have been better to repeat the original process as done by \cite{rizoiu2018debatenight} through the use of 
 100k synthesized cascades, but it required the creation of huge synthesized Adjacency Matrix. Due to limitations in computing, this was not possible.

 \section{Data Analysis}
\label{sec:analysis}
\subsection{Data Captured}
After cascading and re-scraping, the actual size of the top cascades on Twitter was compared with the size we had. This comparison has issues. There is a selection bias in this data as only the top tweets were re-scraped. 
Additionally, we re-scraped metrics from twitter in July 2019, nine months after the collection. Tweets may have been removed, and new retweets may have been made.
However, this analysis is still included for future comparisons. \par

In our 329606 cascades, we had re-scraped the details of 154603 parent
tweets. We compared the total retweet and replies we had (when re-scraped) with the total retweets and replies on Twitter. After summing the total tweet and retweet amount in Twitter 
and comparing it with the sum we had managed to captured, it was found that we had captured 9.64\% of data in each cascade. \par

@@

\subsection{Exploratory Data Analysis}
After the removal of irrelevant data, 6.5M Tweets by 1.7M users remained. Bots made 12.6\% of the Tweets. We removed stopwords to compare the words used by bots and humans. 
We selected words which appear more than 0.03\% times in the dataset (which was 45759 times for humans and 6538 for bots). Wordcloud was created to visualize 
the differences.

\begin{figure}[h!]
    \includegraphics[width=\linewidth]{images/mostly_bots.png}
    \caption{Word used more more often by bots}
    \label{fig:bots_dominant}
\end{figure}

@@

From Table \ref{tab:bot-dominant-words} and Figure \ref{fig:bots_dominant}, we can see that bots dominate humans in the use of conservative words.

\begin{figure}
    \includegraphics[width=\linewidth]{images/mostly_humans.png}
    \caption{Word used more more often by humans}
    \label{fig:humans_dominant}
\end{figure}

@@

Table \ref{tab:humans-dominant-words} and Figure \ref{fig:humans_dominant} shows that humans dominate bots in the use of slang. \par

\begin{figure}
    \includegraphics[width=\linewidth]{images/emotions.png}
    \caption{Histogram of Tweet Sentiment}
    \label{fig:setiment_histogram}
\end{figure}

In Figure \ref{fig:setiment_histogram}, a histogram was created to visualize the difference in Sentiment difference between Human and Bots. 
The figures are very similar and do not show any significant differences.\par

The 329606 cascades in the dataset was analyzed to calculate human-bot interaction. The following results was observed:

@@

  Next, cascades started by bots, and humans were compared with each other. The cascades were grouped by their length to properly visualize the differences. 
  In Figure \ref{fig:further_retweet_bots} we can see that percentage amount of bots was high for smaller bot started cascades. It becomes smaller as the size increases. 
  Opposite was observed for humans in Figure \ref{fig:further_retweet_humans}. It might possibly mean that humans and mostly followed by humans and bots by bots. Both join once a tweet becomes famous.

>>>>

\begin{abstract}
    Use of bots in social media and its consequences has been a subject of great public interest. In this study, we measure the influence of bots during 
    the 2018 US Midterm Elections. Between October 31, 2018 and November 7, 2018 we collected 6.48M tweets and 1.6M profile data from Twitter. We used 
    Machine Learning techniques to separate bot and human accounts and found that 1\% of the
    top bots influenced the creation of 11\% of all the tweets. Additionally, for URLs which were retweeted more than ten times in our dataset, we found that the 
    total influence percentage of bots was double their size. 
    Influence is a number we give to a tweet which measures 
    the expected number of time the tweet gets retweeted over all possible diffusion scenarios. Influence is converted to percentage by using the sum of user wise influence 
    throughout the dataset.
\end{abstract}

@@

Twitter is a social media where short messages get exchanged. Messages shared by a user are shown to
their followers and to followers of those who share the message. These messages are called Tweets. Sharing can be done by retweeting
(sharing) or replying to the tweet. When Alice replies/retweets, a tweet made by Bob, Alice and Bob's followers will see the tweet. In this
way, messages spread. Having massive followers, getting many retweets, getting retweeted by people with massive followers or advertising can make a tweet viral.\par

People use social media, at least in part, to form an opinion about lifestyle, health, politics, and purchase \cite{varol2017early}. 
Due to the power of social media in influencing opinion, various ethical and unethical techniques have been devised to reach a big audience. 
One such unethical technique is the use of bots. Social bots are account controlled wholly or partially by a computer algorithm. 
These bots can generate content and interact with human users often imitating humans \cite{ferrara2016rise}.
A US special counsel investigation found that Twitter accounts of US personas were being operated as a botnet (a network of bots) to amplify content. 
During the 2016 US Presidential Election, they chose candidates to support and oppose \cite{luceri2019red,mueller_investigation}. 
These propaganda bots were found to be high volume, multichannel, rapid, continuous, repetitive, inaccurate, and inconsistent \cite{paul2016russian}. \par

Past studies have explored the political polarization \cite{munger2017don,rizoiu2018debatenight,gruzd2014investigating,bovet2019influence}, origin of bots 
\cite{zannettou2019let,zannettou2019characterizing} and the spread of fake news \cite{vosoughi2018spread,shao2018spread,lazer2018science,bovet2019influence,grinberg2019fake}
on twitter. Previous US Election has been studied before. Most studies focus on 2016 Election \cite{bovet2019influence,rizoiu2018debatenight,bessi2016social,howard2018algorithms,howard2016bots}. \cite{deb2019perils} 
studied the 2018 Midterm Election.\par

Few studies perform a large scale quantitative study of the role played by bots during a significant event. To fill the gaps in the literature, 
we measured the influence of bots during the 2018 US Midterm Election. \par

@@

\begin{itemize}
    \item 10.5\% of the users in the datasets were bots. They made 12.6\% of the tweets and created 22\% of the total influence.
    \item On Average, bots were twice as influential as humans
    \item Top 1\% of the bots influenced the creation of 11\% of all the retweets during this time.
    \item In average a single URL was shared 27.84 times in 1.97 cascades. URL cascade is formed when a Tweet mentions a URL. Any further retweet belong to the same cascade.
    \item For URLs which were shared more than ten times, the influence percentage of bots was double their size.
    \item 81.64\% of all interaction (retweets and replies) took place between humans, and 2.87\% took place between bots. 6.37\% was bot-human, and 9.14\% was human-bot interaction.
    \item Most of the codes and notebooks are shared publicly in https://github.com/warproxxx/2018Midterm to help future researchers and ensures transparency.
\end{itemize}

@@

Different types of bots exist in social media. A simple bot will post predetermined messages at predetermined intervals \cite{haustein2016tweets}. Bots are also used to increase followers
and make an account appear popular \cite{cresci2015fame}. Botnets are a network of bots working together. The word Botnet is also used to refers to a network of compromised computer. However, in this report, 
the word botnet refers to a network of social media bots working together to make an influence and should not be confused with the other definition. Many botnets use a hybrid human/automation 
approach \cite{grimme2018changing}. Botnets have been used to promote spam \cite{ferrara2018measuring}, which seem to shifted to social media due to the effectiveness of spam filter \cite{gao2010detecting,chu2012detecting,ferrara2018measuring}, 
manipulate stocks \cite{ferrara2015manipulation}, manipulate elections \cite{morstatter2018alt}, 
and for various other purposes \cite{abokhodair2015dissecting}.
Botnets were detected as early as 2010 US Midterm Election \cite{mitter2014categorization}. Over time, their use and study have increased. 
Studies have detected and analyzed bots in the 2017 German Federal Election \cite{morstatter2018alt} ,2017 French Presidential Election \cite{ferrara2017disinformation} and during 
various US Elections \cite{mitter2014categorization,bovet2019influence,rizoiu2018debatenight,bessi2016social,howard2018algorithms,howard2016bots,deb2019perils}. In Elections, 
they have been used to support candidates \cite{luceri2019red}, attack people \cite{mueller_investigation} and spread fake news \cite{vosoughi2018spread,grinberg2019fake}. 
Political bots are active beyond the election time. \cite{stewart2018examining} found that Russian bots infiltrated both right and left-leaning communities and spread different narratives 
\cite{mueller_investigation}. \par

In Twitter, bots share content 
in multiple channel \cite{paul2016russian}. This activity is in line with literature which shows that information from multiple sources appears more trustworthy than a single one \cite{harkins1981multiple}. 
 Bots attack people. Attacking trustworthiness has been shown to diminish the credibility of the Original Poster \cite{pornpitakpan2004persuasiveness}. 
Bots are continuous, repetitive, but most of the times, they are also far from reality \cite{paul2016russian}. The high volume seems to ensure early exposure and \cite{petty1994think} shows that early
 first impression is more likely to be accepted by the brain. Due to a phenomenon called Sleeper Effect, where information gets disassociated from the source while remembering \cite{underwood1998memory,paul2016russian}
 low credibility sources can have persuasive power to unbiased or neutral people. As a boost to sleeper effect, people are 31\% more likely to remember what they see on twitter, 
 compared to the normal web \cite{twitter_remember}. \par

\subsection{Bot Detection}
Most studies \cite{rizoiu2018debatenight,yang2019arming,shao2018spread} use Botometer\cite{davis2016botornot,yang2019arming} for bot detection.
Botometer's Machine Learning algorithm provides an account-level bot classification. Botometer classifies account partially and completely automatized as bots. 
Botometer has some false positives. Its detection of organizational accounts as bots have been highly criticized \cite{varol2017early} \cite{botometer_tweet}. However, even with its false positive, 
Botometer has no transparent competitor and 
remains the most accurate bot detection system in academia. \par

@@

\cite{rizoiu2018debatenight} found that social bots were 2.5 times more influential than humans. They introduce a scalable algorithm for estimating user influence in retweet cascades. 
For each tweet in an information cascade, it uses time of post and the number of followers to determine the probability of whether a tweet is a retweet of another. They then find influence which can be used 
to find out how many users the tweet possibly influenced to retweet. It was tested successfully in artificial social media. Other studies have attempted to measure influence before this. 
\cite{weng2010twitterrank} used eigenvector centrality of the connection to measure influence, but it is not scalable into big cascades. There are other methods like 
\cite{rodriguez2011uncovering,cho2013latent,linderman2014discovering} but they either have scaling issue or require full diffusion graph, which Twitter does not provide.

\section{Methods}
\label{sec:method}
\subsection{Data Collection}
Twenty-nine manually identified keywords, and names of all 53 House candidates and 27 Senate candidates were used to collect live data from Twitter. The manually defined keywords are in 
Table \ref{tab:manual-keywords}

@@

Twitter provides Streaming API, Search API, and a premium Firehose API to provide access to tweets. Streaming API provides limited access to live data (thus also called 1\% API)
 while Firehose API
provides complete data. However, due to the associated costs, Firehose API was not an option in this study. \cite{morstatter2013sample} compares the Streaming API with Firehose API.
They found that Streaming API was nearly as good as a random sample of Firehose API when the dataset was large enough. Although the 1\% API was not as good as a 1\% random sample from the Firehose API in 
all of their tests, it estimated the top hashtags correctly when the data was large enough. It recommended the creation of specific parameters and using a large sample. 
Due to the possible issues in Streaming API, \cite{bessi2016social} recommends using Twitter Search API.

\par In this study, we use Streaming API with many keywords as suggested by \cite{morstatter2013sample} to collect a large dataset. While collecting data, information about the tweeter and the tweet was collected.

We collected following information about a tweet:
\begin{enumerate}[label=\textbf{\arabic*}]
    \item \textbf{Timestamp:} UTC Timestamp in which the post was made
    \item \textbf{ID:} Post ID provided by twitter
    \item \textbf{Text:} Text in the tweet and the parent tweet if the tweet is a reply
    \item \textbf{User:} Username of the tweeter/retweeter/replier
    \item \textbf{Replies:} The number of replies the tweet has received. During live collection, this value is zero. We collect them during recollection of top tweets.
    \item \textbf{Retweets:} The number of retweets the tweet has received. During live collection, this value is zero. We collect them during recollection of top tweets.
    \item \textbf{Likes:} The number of likes the tweet has received. During live collection, this value is zero. We collect them during recollection of top tweets.
    \item \textbf{Reply To ID:} The ID of the parents tweet if this is retweet or a reply.
    \item \textbf{Response Type:} Either Tweet or Retweet or Reply
\end{enumerate}
\bigskip

We collected the following profile information from the Twitter API:
\begin{enumerate}[label=\textbf{\arabic*}]
    \item \textbf{Username:} Username of the user who posted the tweet
    \item \textbf{Location:} Binary if geolocation is enabled
    \item \textbf{Is Verified:} A binary if the profile has been verified
    \item \textbf{Total Tweets:} Total number of tweets created by the user
    \item \textbf{Total Following:} Total accounts the user is following
    \item \textbf{Total Followers:} Total Followers
    \item \textbf{Total Listed:} Number of times a user has been listed
    \item \textbf{Total Status:} Total Status
    \item \textbf{Total Likes:} Total likes the user has received
    \item \textbf{Has Background:}  Binary if an account has a background
    \item \textbf{Is Protected:}  Binary if an account is protected
    \item \textbf{Profile Modified:}  Binary if a profile has been modified
\end{enumerate}
\bigskip
\cite{kudugunta2018deep} and \cite{ferrara2017disinformation} use the same parameters in their machine learning architecture. \par

We collected Tweets from October 31, 2018 (1 AM UTC) to November 7, 2018 (midnight UTC). On July 18, 2019, we used Twitter API to rescrape the top 100k most retweeted which did not 
exist in our dataset.

\subsection{Data Processing}
First, we manually removed irrelevant tweets. We created keywords find and remove these irrelevant data.\par

We used SentiStrength and VADER Sentiment analysis in the Tweets. SentiStrength \cite{thelwall2010sentiment} is used to annotate short, informal tweet like texts \cite{bessi2016social}. 
It can capture positive and negative emotions at an accuracy of 60.6\% and 72.8\%. SentiStrength gives a positive and negative score between 0 and 4. The positive sentiment is subtracted 
from the negative like in \cite{bessi2016social} to get a whole number. VADER is a rule-based sentiment classification designed for use in social media data. VADER was found to outperform 
individual human raters with an F1 accuracy of 0.96 compared to 0.84 for humans \cite{hutto2014vader}.

@@

\begin{itemize}
    \item \cite{cresci2017paradigm} provided a labeled account level bot or not data with relevant features that included the ones we collected.
    We used these features to train a Random Forest Classifier and detect bots. 
    \cite{cresci2017paradigm}'s data is old, and bots evolve quickly. So further processing was required.
    \item We used the trained model on our dataset, which contains the same features, to select 40k highly probable bots.
    \item We selected 40k highly probable bots and 32k random accounts from our dataset and performed bot detection through Botometer. Botometer returns CAP score and bot score. CAP score denotes the probability of being a bot. 
    Previous studies have used bot score of 0.5 to determine bots. CAP score of 0.3 is equivalent to the score 0.5 used by researchers in these earlier studies. \cite{deb2019perils} used a threshold of 0.3 after the 
    introduction of new changes as mentioned in \cite{yang2019arming}.
    \item Grid search was performed with thresholds of 0.1, 0.2, and 0.3 for humans, 0.3, 0.5, 0.7, and 0.8 for bots and Random Forest, Neural Network
     and SMOTENN as possible ML algorithms to determine the best parameters in the training and test set
    \item From Grid search, we determined that 0.5 was the optimal bot threshold, 0.3 the optimal human threshold and Adaboost with SMOTENN preprocessing the optimal ML algorithm. Using this threshold, we selected 513 bots whose CAP score was higher than 0.5  and 
    27k human accounts whose CAP score was smaller than 0.3 for training and testing our data. We created a training-test split at a ratio of 0.9. We used the remaining 44k values that lied between 0.5 and 0.3 as a test set,
    which we name "Broader Test set" to see the models performance in thresholds that were not selected. 
\end{itemize}

We obtained an accuracy of 85\% with AUC of 0.85 in the training set, 87\% accuracy with 
an AUC of 0.8 in the test set and accuracy of 85\% with an AUC of 0.8 in the broader test set when compared with Botometers prediction. This model detected 10.5\% of the accounts as 
bots. The accuracy was lesser than in \cite{kudugunta2018deep}. It may be due to the evolution of bots, or because we compared with Botometer instead of manually labeling the training set like in  \cite{kudugunta2018deep}.

\subsection{Cascading and Influence Detection}
A tweet cascade starts when a Tweet is made. Any retweet or response which involve that Tweet belongs to that cascade. If Bob retweets Alice's tweet, and Eve retweets Bob's retweet, the tweets
made by Alice, Bob, and Eve will belong in the same cascade. The Twitter API will show Eve and Bob as a direct descendant of Alice without any mention that Eve retweeted Bob's retweet. \par

\cite{du2013scalable} defines influence as the average number of users who get in contact with the content created by a user \textit{u}.
 However, Twitter does not provide the diffusion graph and the number of people reached. \cite{rizoiu2018debatenight} define the influence of a user over a retweet cascade as "the expected number of time the tweet is
 retweeted – direct retweets or descendants in a diffusion scenario – over all possible diffusion scenarios  associated." They define the influence of a user as the sum of the influence of tweets
 authored by a user. We use their algorithm and definition. \cite{rizoiu2018debatenight} uses time and 
 number of followers to estimate a user influence. Using this mechanism, a high influence score can be provided to highly connected users who never start diffusions and 
 to active retweeters with little followership. \cite{rizoiu2018debatenight} tested their algorithm on an artificial social network with 1000 users and found that the influence calculated had a Spearman 
 correlation coefficient of 0.88 with the actual influence. \par

@@

 \cite{rizoiu2018debatenight} tested the validation of their algorithm when they had access to full data. As we used streaming API, we did not. So random cascades were synthesized using the following algorithm 
 to test the validity:

@@

Then we converted the raw influence score to a user wise percentage. A correlation of 0.96 existed between the full dataset and the 1\% sample. We used this method 
 instead of \cite{rizoiu2018debatenight}'s original one due to limitations in computing power.

 Influence percentage mean. This way, it measures the role of a user in creating new tweets. More the influence percentage, more tweets in the dataset were created by that user/user group. In 
a cascade, the original tweet is the parent tweet.

 \section{Data Analysis}
\label{sec:analysis}
\subsection{Data Captured}
After cascading and re-scraping the missing tweets, we compared the actual size of the cascades with the size we had captured using the Twitter API. This comparison has issues. 
There is a selection bias in this data as we selected cascades which were longer than 2 Tweets. Additionally, we re-scraped metrics (likes, replies and retweets) from twitter in July 2019, nine months after
 the collection. By this time, addition and removal of tweets will take place. However, we still include this analysis to enable future comparison with other researchers. \par

In our 329606 cascades, we had re-scraped the details of 154603 parent
tweets. We compared the total retweet and replies we had (when re-scraped) with the total retweets and replies on Twitter. After we added the total tweet and retweet amount in Twitter  
and compared it with the amount we captured, we found that we had captured 9.64\% of data in each cascade. \par

@@

\subsection{Exploratory Data Analysis}
After the removal of irrelevant data, 6.5M Tweets by 1.7M users remained. Bots made 12.6\% of the Tweets. We removed stopwords to compare the words used by bots and humans. 
We selected words which appear more than 0.03\% times in the dataset (which was 45759 times for humans and 6538 for bots). The dominant words are in 
\ref{tab:bot-dominant-words} and  \ref{tab:humans-dominant-words}. From Table \ref{tab:bot-dominant-words}, we can see that bots dominate humans in the use of conservative words. 
Table \ref{tab:humans-dominant-words} shows that humans dominate bots in the use of slang. \par

@@

@@

We did not find any significant difference between the sentiment of humans and bots. Then we analyzed the 329606 cascades in the dataset to calculate human-bot interaction. Out of all interactions, we made the following observations:

@@

Next, we compared the cascades started by bots and humans. The cascades were grouped by their length to visualize the differences accurately. 
In Figure \ref{fig:further_retweet_bots}, we can see that the percentage of bots was high for smaller cascades started by bots. As the size of Cascade increases, the percentage of bots decreases. 
We made the opposite observation for humans in Figure \ref{fig:further_retweet_humans}. If a tweet receives few audience, they are most likely the followers. This means that bots are mostly
followed by bots and humans by humans. Thus, this finding goes in line with those made by earlier research which shows that bots and humans create a follower cluster dominated by their type.

To further analyze this, in Figure \ref{fig:humans_bots_percentage}, we look at the successful tweets and the early retweets. The early audience of a Tweet are the follower of the Original Poster (OP). If a Tweet becomes famous,
it goes beyond the followers. From the figure, we can see that the number of bots is high in first one percent of the retweets/replies for successful tweets in case of bots and is high for humans in a decreasing amount. This analysis provides further support to the assertion made in the previous paragraph.
